---
title: "Using Meteorological Data to Predict Solar Power Generation"
author: "by [Cormac Nolan](https://github.com/cormac85/) - `r format(Sys.time(), '%d %B %Y')`"
output: 
  html_document:
    code_folding: "hide"
    css: style.css
    includes: 
      after_body: footer.html
      in_header: header.html
editor_options: 
  chunk_output_type: console
---

# {.tabset .tabset-fade .tabset-pills}
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Introduction
Can I use the solar irradiation data from a nearby weather station to my house
to predict the solar power generated by the rooftop solar panels I have?
```{r libraries, message = FALSE, warning = FALSE}
library(dplyr)
library(curl)
library(jsonlite)
library(ggplot2)
```


```{r functions}

```

I've imported data directly from the Met Eireann API that shows monthly
historical data for a nearby weather station at Dunsany, in Co. Meath. I've
also transcribed a dataset from the logging app for my solar inverter which
details how much power was generated per month in kWh.
```{r import, cache=TRUE}
report <- jsonlite::fromJSON("https://prodapi.metweb.ie/monthly-data/Dunsany%20(Grange)")

monthly_solar_generation <- dplyr::tibble(
  year = c(rep("2021", 12), rep("2020", 12), rep("2019", 12), rep("2018", 12)),
  month = rep(1:12, 4),
  solar_power_generation = c(
    c(85, 157.5, 44.5, rep(NA, 9)),
    c(rep(NA, 5), 358.4, 394.9, 314.3, 332.8, 235, 119.6, 38.7),
    rep(NA, 12),
    rep(NA, 12)
    )
  )

```

## Cleaning
We extract and clean up the Met Eireann solar irradition data and then combine
it with the solar power generation dataset above. We filtered it to when we 
have complete data.
```{r raw exploration}

dunsany_solar_radiation <- 
  report$solar_radiation$report %>% 
  purrr::map_df(as_tibble) %>% 
  mutate(year=c("2021", "2020", "2019", "2018", "mean")) %>% 
  select(-annual) %>% 
  filter(year != "mean") %>% 
  tidyr::pivot_longer(january:december, names_to = "month", values_to = "solar_irradiation_jpercm2") %>% 
  mutate(solar_irradiation_jpercm2 = dplyr::na_if(solar_irradiation_jpercm2, ""),
         solar_irradiation_jpercm2 = dplyr::na_if(solar_irradiation_jpercm2, "n/a"),
         month = rep(1:12, 4),
         solar_irradiation_jpercm2 = as.numeric(solar_irradiation_jpercm2))

dunsany_solar_radiation <- 
  dunsany_solar_radiation %>% 
  inner_join(monthly_solar_generation, by = c("year", "month"))

reduced_dunsany_solar_radiation <- 
  dunsany_solar_radiation%>% 
  filter(!is.na(solar_irradiation_jpercm2),
         !is.na(solar_power_generation))

reduced_dunsany_solar_radiation
```

## Exploration
### Correlation
We see a very high correlation between the solar power data and the irradiation
data, which is expected but good to see in practice.
```{r correlation 1}
cor(reduced_dunsany_solar_radiation$solar_irradiation_jpercm2,
    reduced_dunsany_solar_radiation$solar_power_generation)

```

The correlation appears statistically significant.
```{r exploration 2}
cor.test(reduced_dunsany_solar_radiation$solar_irradiation_jpercm2,
    reduced_dunsany_solar_radiation$solar_power_generation)
```

### Plots
```{r}
reduced_dunsany_solar_radiation <- 
  reduced_dunsany_solar_radiation %>% 
  mutate(month = sprintf("%02d", as.numeric(month))) %>% 
  mutate(year = as.integer(year)) %>% 
  tidyr::unite("year_month", year, month, sep="-", remove=FALSE) %>% 
  mutate(
    normalised_irradiation = solar_irradiation_jpercm2 / max(solar_irradiation_jpercm2),
    normalised_solar_power = solar_power_generation / max(solar_power_generation)
  )

reduced_dunsany_solar_radiation %>% 

  select(year_month, normalised_irradiation, normalised_solar_power) %>% 
  tidyr::pivot_longer(cols = normalised_irradiation:normalised_solar_power,
                      names_to = "solar_measurement",
                      values_to = "normalised_measurement_value") %>% 
  ggplot2::ggplot(
    aes(year_month,
        normalised_measurement_value, 
        group=solar_measurement,
        colour=solar_measurement)
  ) +
  geom_line() +
  scale_color_brewer(palette = 2, type = "qual") +
  theme_minimal()
  

```


## Modelling
### Simple Linear Model
```{r model 1}
solar_model <- lm(
  solar_power_generation ~ solar_irradiation_jpercm2, 
  data = reduced_dunsany_solar_radiation
)

```


```{r model 2}
predict(solar_model, data.frame(solar_irradiation_jpercm2 = 15000))
```

### Transformations
```{r model 3 }

```


```{r}

```


```{r}

```


```{r}

```
